---
title: "ACM poster 2017"
output: html_notebook
author: "Daniela S. Gil"
---
# Intro 

In this report we are focusing on the most frequented edges, all having more than 170 sessions along 
April 2016. All of this edges are around FEUP. 

The data frame to use is: df_superhotedges_april16pt.
    It was created with the df df_speed_top which brings directly all points 
    of the top edges from the database. 
    The following filters were applied (more info: PreAnalysis_topedges.Rmd)
      *Week days
      *Hour between 8hr and 20h
      
    Given a total of 9 variables, 9 edges, 1369 sessions and 190201 points. 
    Half of the the sessions involve just one edge.  
    1011 pass by Frei
    150 sessions have gaps over 5min 
    394 sessions have gaps over 2min 


```{r echo=FALSE}
#Creating a df with all the points in the sessions that pass by SHE 
df_superhotedges_april16pt_all <- df_edges_april16pt %>% 
  filter(session_id %in% df_superhotsessions_april16pt)

```

```{r echo=FALSE}

df_superhotedges_april16pt  %>%
  merge(df_edges_april16pt, using = session_id) %>% 
  group_by(session_id) %>%
  summarise(n = n_distinct(way_id)) %>% 
  ggplot(aes(n)) +
  geom_bar(aes(fill=factor(n)))

```

```{r echo=FALSE}
df_superhotedges_april16pt  %>%
  group_by(session_id) %>% 
  summarise(n = n_distinct(segment_id)) %>% 
  ggplot(aes(n)) +
  geom_bar(aes(fill=factor(n)))

```

# Map with number of sessions
```{r echo=FALSE}

########### Selecting edges to map #########

# All edges (in Porto) 
df_sessions_edges_april16pt <- df_intersession_april16pt %>% 
  group_by(way_id) %>% 
  summarise(sessions = n_distinct(session_id)) %>% 
  merge(y= df_points_edge_osm_april16pt, using = way_id) 

df_osm_edge <- df_sessions_edges_april16pt %>% 
  select(way_id, sessions)

## Discretization based on number of sessions 
df_sessions_edges_april16pt <- df_sessions_edges_april16pt %>% 
  mutate(category = case_when(sessions > 155 ~ "SHE",
                              sessions >= 50 & sessions <155 ~ "medium",
                              sessions > 25 & sessions < 50 ~ "low"))


# Hot edges 
df_osm_edge <- df_sessions_edges_april16pt  %>% 
  subset(way_id %in% df_osm_edge_ids$way_id) %>% 
  select(way_id, sessions)

######### List before passing to map. #########
list_osm_edge <- df_osm_edge[, 1]

########## Map in April2016.Rmd #########
```

```{r}
ms_1 <- m
ms_1
m
```


```{r}
# Map with all points with nº sessions
#     green: <50, 50< yellow <155, red > 155 

#ms_3 <- m 
ms_3
```

# Intersession time

```{r echo=FALSE}
 df_intersession_april16pt %>% 
  filter(intersession_time > 0 ) %>% 
  merge(df_sessions_edges_april16pt, using = way_id) %>% 
  ggplot( aes(x=intersession_time, color = category))+
  scale_x_log10(limits = c(1e0,1e4), breaks = c(10,30,60,1000)) +
  geom_line(aes(y = 1 - ..y..), stat='ecdf') +
  scale_color_manual(values = c("grey","blue","red","green"),
                     breaks = c("SHE", "medium", "low","NA"),
                     labels = c("High", "Medium", "Low","NA")) +
  theme(legend.title = element_blank(),
        legend.text = element_text( size = 20, face = "bold"),
        legend.position = c(0.8,0.8),
        axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=20),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=20),
        panel.background = element_rect(fill = 'white', colour = 'black')) +
    xlab("Intersession time (min)") + 
    ylab("P(X>=x)") +
  geom_vline(xintercept =  c(15,30,45,60), color ="dark grey") 
  
```


```{r echo=FALSE}
 df_intersession_april16pt %>% 
  filter(intersession_time > 0 ) %>% 
  merge(df_sessions_edges_april16pt, using = way_id) %>% 
  ggplot( aes(x=intersession_time, color = category))+
  scale_x_continuous(breaks) +
  geom_line(aes(y = 1 - ..y..), stat='ecdf') +
  scale_color_manual(values = c("grey","blue","red","green"),
                     breaks = c("SHE", "medium", "low","NA"),
                     labels = c("High", "Medium", "Low","NA")) +
  theme(legend.title = element_blank(),
        legend.text = element_text( size = 20, face = "bold"),
        legend.position = c(0.8,0.8),
        axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=20),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=20),
        panel.background = element_rect(fill = 'white', colour = 'black')) +
    xlab("Intersession time (min)") + 
    ylab("P(X>=x)") +
  geom_vline(xintercept =  c(15,30,45,60), color ="dark grey")


```


# Interpoint time 

```{r echo=FALSE}
#df_superhotedges_april16pt$interpoint_time <- NULL

#df_superhotedges_april16pt <- 
  
df_superhotedges_april16pt <- df_superhotedges_april16pt  %>%
  group_by(session_id) %>% 
  arrange(time) %>% 
  mutate(interpoint_time = c(0,as.numeric(diff(time), units="mins"))) 
```

```{r}
boxplot(e$interpoint_time, ylim = c(0,50))
```

```{r}
#Different sessions with interpoint > x
e %>% 
  filter(interpoint_time >= 2) %>% 
  count
```

```{r}
df_superhotedges_april16pt %>% 
  filter(interpoint_time > 1) %>% 
  ggplot(aes(factor(way_id),interpoint_time)) +
  geom_boxplot() + 
  coord_cartesian(ylim= c(0,50))

```

```{r}
df_superhotedges_april16pt %>% 
  group_by(session_id) %>% 
  summarise(n = n_distinct(session_id))
  

  filter(way_id != 37141967) %>% 
  fgroup_by(session_id) %>% 
  summarise(speed = quantile(speed,0.9)) %>% 
  filter(speed == 0)
```


# Temporal

```{r echo=FALSE}
df_edges_april16pt %>%
  mutate( category = case_when(way_id %in% df_osm_edge_ids$way_id ~ "SHE", TRUE ~ "ALL")) %>% 
  group_by(hour =hour(time), category) %>% 
  summarise(sessions = n_distinct(session_id)) %>% 
  ggplot(aes(hour, sessions, color = category)) + 
  geom_line()+
  scale_x_continuous(breaks = seq(0,23,1))
```

```{r echo=FALSE}
df_edges_april16pt %>%
  merge(df_sessions_edges_april16pt, using = way_id) %>% 
  group_by(hour =hour(time), category) %>% 
  summarise(sessions = n_distinct(session_id)) %>% 
  ggplot(aes(hour, sessions, color = category)) + 
  geom_line()+
  scale_x_continuous(breaks = seq(0,23,1))
```

```{r}
df_sessions_allmodes_april16pt %>% 
  group_by(hour = hour(time_start)) %>% 
  summarise(n_sessions = n_distinct(session_id)) %>% 
  plot_ly(x=~factor(hour), y=~n_sessions, type = 'scatter', mode = 'lines')

df_sessions_allmodes_april16pt <- df_sessions_allmodes_april16pt %>% 
  group_by(travelmode) %>% 
  mutate(total_sessions = n_distinct(session_id))

df_sessions_allmodes_april16pt %>% 
  filter(travelmode != "NA") %>% 
  group_by(hour = hour(time_start), travelmode, total_sessions) %>% 
  summarise(n_sessions = n_distinct(session_id)) %>% 
  mutate(freq = n_sessions/total_sessions *100) %>% 
  ggplot(aes(x = hour,y = freq, color=travelmode)) +
  geom_line() + 
  ylab("Frequency(%)") +
  xlab("Hour") +
  scale_x_continuous(breaks = seq(0,23,1)) + 
  scale_color_discrete(name = "Travelmode") +
  theme(axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=16),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=16),
        legend.title = element_text("Street", size=16, face="bold"), 
        legend.text = element_text(size=16, face="bold"),
        panel.background = element_rect(fill = 'white', colour = 'black'),
        legend.position = c(0.2,0.7))

```

```{r}
df_sessions_allmodes %>% 
  filter(travelmode != "NA") %>% 
  group_by(hour = hour(time_start), travelmode) %>% 
  summarise(n_sessions = n_distinct(session_id)) %>% 
  ggplot(aes(x = hour,y = n_sessions, color=travelmode)) +
  geom_line() + 
  scale_x_continuous(breaks = seq(0,23,1)) + 
  theme_light()

```


# Origin-Destination 

```{r}
df_st_hotedges_april16pt <- df_hotedges_april16pt[,-2] %>% 
  group_by(session_id) %>%
  slice(which.max(time)) %>% 
  mutate(end = case_when((lat > 41.175645 & lat < 41.180481 & lon > -8.607200 & lon < -8.586093) ~ "FEUP",  
                           TRUE ~ "OUT"))  %>% 
  rename(c("way_id"="way_id_end","segment_id"="segment_id_end", "seconds"="seconds_end","lat" = "lat_end","lon" = "lon_end","speed" = "speed_end", "time"="time_end")) 
  
df_st_hotedges_april16pt <- df_hotedges_april16pt %>% 
  group_by(session_id) %>%
  slice(which.min(time)) %>% 
  mutate(start = case_when((lat > 41.175645 & lat < 41.180481 & lon > -8.607200 & lon < -8.586093) ~ "FEUP",  
                           TRUE ~ "OUT")) %>% 
  rename(c("way_id"="way_id_start","segment_id"="segment_id_start", "seconds"="seconds_start","lat" = "lat_start","lon" = "lon_start","speed" = "speed_start", "time"="time_start")) %>% 
  merge(y = df_st_hotedges_april16pt, using = way_id)


df_st_hotedges_april16pt <- df_st_hotedges_april16pt %>% 
  mutate(class_trip = case_when((start == "FEUP" & end == "OUT") ~ 1,
                    (start == "OUT" & end == "FEUP") ~ 2,
                    (start == "OUT" & end == "OUT") ~ 3,
                    (start == "FEUP" & end == "FEUP") ~ 4)) 

df_st_hotedges_april16pt$class_trip <- as.factor(df_st_hotedges_april16pt$class_trip)
df_st_hotedges_april16pt$end <- as.factor(df_st_hotedges_april16pt$end)
df_st_hotedges_april16pt$start <- as.factor(df_st_hotedges_april16pt$start)
str(df_st_hotedges_april16pt)

#All  
df_st_hotedges_april16pt %>% 
  mutate(class_day = case_when(wday(time_start)!= 1 &  wday(time_start)!= 7 ~ "Weekday",
                   TRUE ~ "Weekend")) %>% 
  ggplot(aes(x=hour(time_start))) + 
  geom_bar(aes(fill = class_trip )) + 
  scale_fill_hue(breaks = c("1", "2", "3","4"),
                 labels = c("From campus", "To campus", "Outside campus","Within campus"))+
   theme(legend.title = element_blank(),
        legend.text = element_text( size = 20, face = "bold"),
        legend.position = c(0.75,0.8),
        axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=16),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=16),
        panel.background = element_rect(fill = 'white', colour = 'black')) +
  xlab("Hour") + 
  ylab("# of sessions") +
  facet_grid(~as.factor(class_day)) + 
       theme(strip.text.x = element_text(size = 20))

```

```{r echo=FALSE}
# Week days  
st_wd <- df_st_hotedges_april16pt %>% 
  filter(wday(time_start)!= 1,  wday(time_start)!= 7) %>% 
  ggplot(aes(x=hour(time_start))) + 
  geom_bar(aes(fill = class_trip ))+
  theme(legend.title = element_blank(),
        legend.text = element_text( size = 20, face = "bold"),
        legend.position = c(0.2,0.8),
        axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=16),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=16),
        panel.background = element_rect(fill = 'white', colour = 'black')) +
  xlab("Hour") + 
  ylab("Nº sessions")

```

```{r echo=FALSE}
# Weekend
st_we <- df_st_hotedges_april16pt %>% 
  filter(wday(time_start)== 1 | wday(time_start)== 7) %>% 
  ggplot(aes(x=hour(time_start))) + 
  geom_bar(aes(fill = class_trip )) +
  theme(legend.title = element_blank(),
        legend.text = element_text( size = 20, face = "bold"),
        legend.position = c(0.3,0.8),
        axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=16),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=16),
        panel.background = element_rect(fill = 'white', colour = 'black')) +
  xlab("Hour") + 
  ylab("Nº sessions")

multiplot(st_wd, st_we, cols = 2)
```


# External Factors(parking, etc)

```{r}
df_superhotedges_april16pt %>% 
  filter(way_id == 37141967, day(time) == 20)  %>% 
  group_by(way_id, session_id, segment_id) %>% 
  summarise(iqr = IQR(track)) 

df_superhotedges_april16pt %>% 
  filter((way_id == 37141967 & session_id == 1806847192))  %>% 
  plot_ly( x= ~time, y= ~track,color = ~factor(segment_id), type ="scatter")

```

## Frei  

```{r}
df_superhotedges_april16pt %>% 
  filter(way_id == 37141967)  %>% 
  group_by(way_id, session_id) %>% 
  summarise(iqr = IQR(track), m_speed = median(speed), avg_speed = mean(speed)) %>% 
  filter(m_speed < avg_speed , iqr > 100)

```

```{r}
f3 <- df_superhotedges_april16pt %>% 
  filter( session_id == 1577)  %>% 
  plot_ly( x= ~time, y= ~track,color = ~factor(segment_id), type ="scatter", mode = "lines")

f4 <- df_superhotedges_april16pt %>% 
  filter( session_id == 1577)  %>% 
  plot_ly( x= ~time, y= ~speed,color = ~factor(way_id), type ="scatter", mode ="lines") 

subplot(f3,f4, nrows = 2)

subplot(f1,f2, nrows = 2)
1577
```

## Manuel Silva

```{r}
df_superhotedges_april16pt %>% 
  filter(way_id == 35078500)  %>% 
  group_by(way_id, session_id) %>% 
  summarise(iqr = IQR(track), m_speed = median(speed), avg_speed = mean(speed)) %>% 
  filter(m_speed < avg_speed , iqr > 100)

```

```{r}
 df_superhotedges_april16pt %>% 
  filter((way_id == 35078500 & session_id == 19802))  %>% 
  plot_ly( x= ~time, y= ~track,color = ~factor(segment_id), type ="scatter", mode = "lines")
```


```{r}
df_superhotedges_april16pt %>% 
  filter(way_id == 35078500 & session_id == 19802)  %>% 
  plot_ly( x= ~time, y= ~speed,color = ~factor(way_id), type ="scatter", mode ="lines") 
```

## Density
```{r}
df_superhotedges_april16pt %>% 
  filter( way_id == 35078500 | way_id == 37141967 | way_id == 36979228) %>% 
  ggplot(aes(x= track)) +
  geom_density(aes(group=factor(way_id), color=factor(way_id))) +
  xlab("Grades(º)") +
  ylab("Density") %>% 
  geom_vline(xintercept = c(90,270),color = "grey") + 
  scale_color_discrete(name = "Street",
                       breaks = c("35078500", "37141967", "36979228"),
                      labels=c("1", "2", "3")) +
  theme(axis.title.x = element_text(face="bold", size=20),
        axis.text.x  = element_text(size=16),
        axis.title.y = element_text(face="bold", size=20),
        axis.text.y  = element_text(size=16),
        legend.title = element_text("Street", size=16, face="bold"), 
        legend.text = element_text(size=16, face="bold"),
        panel.background = element_rect(fill = 'white', colour = 'black'),
        legend.position = c(0.1,0.8))
  
  
```

 
## Rascunhos 

```{r}
df_superhotedges_april16pt %>% 
  filter((way_id == 37141967 & session_id == 47419))  %>% 
  merge(df_osm_edge_ids, using =way_id) %>% 
  plot_ly( x= ~time, y= ~track,color = ~factor(segment_id), type ="scatter")

```

```{r}
df_superhotedges_april16pt %>% 
  filter(( session_id == 14249))  %>% 
  plot_ly( x= ~time, y= ~speed,color = ~factor(way_id), type ="scatter") 

long_sessions <- c(11712,19904,11273,85840, 832,3827, 85057, 48182)
```


```{r}
df_hotedges_april16pt %>% 
  group_by(session_id) %>% 
  summarise(n = n_distinct(way_id)) %>% 
  filter(n > 5) %>% 
  arrange((n))

f <- df_edges_april16pt %>% 
  filter(session_id == 35307)


write.csv(f,"f.csv")
```

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( way_id == 35078500 ) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  
```





