---
title: "Sense My FEUP - Mobility forecasting based on GPS location tracking"
author: "Daniela S. Gil"
output: html_notebook
---

# PART I: Base analysis.  

## Package installation. 
```{r echo=FALSE, eval=FALSE}
#Basics. 

if(! "ggplot2" %in% rownames(installed.packages())){
  install.packages("ggplot2")
}

if(! "dplyr" %in% rownames(installed.packages())){
  install.packages("dplyr")
}

if(! "tidyr" %in% rownames(installed.packages())){
  install.packages("tidyr")
}

if(! "xlsx" %in% rownames(installed.packages())){
  install.packages("xlsx")
}

if(! "magrittr" %in% rownames(installed.packages())){
  install.packages("magrittr")
}

if(! "gridExtra" %in% rownames(installed.packages())){
  install.packages("gridExtra")
}

if(! "plotly" %in% rownames(installed.packages())){
  install.packages("plotly")
}

library(gridExtra)
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
#library(xlsx)
#library(magrittr)
#library(lattice)
#library(latticeExtra)
```

```{r echo=FALSE, eval=FALSE}
#connection to database.

if(! "DBI" %in% rownames(installed.packages())){
  install.packages("DBI")
}
if(! "RPostgreSQL" %in% rownames(installed.packages())){
  install.packages("RPostgreSQL")
}

library(DBI)
library(RPostgreSQL)
```

```{r echo=FALSE, eval=FALSE}
#Mapping. 

if(! "leaflet" %in% rownames(installed.packages())){
  install.packages("leaflet",dependencies = TRUE)
}


if(! "sp" %in% rownames(installed.packages())){
  install.packages("sp")
  library(sp)
}

if(! "mapview" %in% rownames(installed.packages())){
  install.packages("mapview",dependencies = TRUE)
  library(mapview)
}

library(htmlwidgets)
library(mapview)
library(leaflet)
```

```{r echo=FALSE, eval=FALSE}
#Time management.

if(! "lubridate" %in% rownames(installed.packages())){
  install.packages("lubridate")
}

library(lubridate)
```

```{r echo=FALSE, eval=FALSE}
#Write Excel
library(xlsx)
library(rJava)
library(xlsxjars)
```

## Connecting the database.
```{r echo=FALSE, eval=FALSE}
# Connecting Sense My FEUP - DS

drv <- dbDriver("PostgreSQL")
pwd <- "psicopatas"

#Database to connect 
db <- "sensemyfeup"
#db <- sensemycity

con <-dbConnect(drv,host='localhost',port='15432',dbname=db,user='danielasocas',
password=pwd)
```

```{r echo=FALSE, eval=FALSE}
# Connecting OpenStreetMap - Marcio Fontes
pw <- {
  "y7dWwByZLWso"
}

drv <- dbDriver("PostgreSQL")

con_osm <- dbConnect(drv, dbname = "openstreetmap",
                     host = "localhost", port = 15432,
                     user = "marciofontes", password = pw)
```

## General EDA.
### Getting data. 
```{r}
df_segments_length <- data.frame()
df_segments_length_tm <- dbGetQuery(con, "SELECT session_id, SUM(length) as distance, travelmode
                                      FROM danielasocas.segments_travelmode as segments 
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'
                                      GROUP BY session_id, travelmode")

df_segments_length <- dbGetQuery(con, "SELECT session_id, SUM(length) as distance
                                      FROM danielasocas.segments_travelmode as segments 
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'
                                      GROUP BY session_id")
```

```{r}
df_segments <- subset(df_segments_length, distance < IQR(distance,0.75) )

ggplot(df_segments, aes(distance)) +
  geom_histogram(bins = 10)

ggplot(df_segments_length, aes(travelmode, distance, color = travelmode)) +
  geom_boxplot()

ggplot(df_segments, aes(distance, color = travelmode)) +
  geom_histogram(binwidth = 1)

ggplot(df_segments, aes(travelmode, distance, color = travelmode)) +
  geom_boxplot()
```


```{r}
df_segments_timestamp <- data.frame()
df_segments_timestamp <- dbGetQuery(con, "SELECT session_id, time_start,time_end, length, travelmode
                                      FROM danielasocas.segments_travelmode as segments
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'")
```


```{r}

```

## Getting data per edge.

### POINTS per edge. 
Getting the points filtered by only segments with travelmode = car or bus.

```{r echo=FALSE, eval=FALSE}
#Obtaining the points through the session and the seconds of the whole segment.  

df_points_edge_osm_cars <- data.frame()
df_points_edge_osm_cars <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_travelmode as segments 
                                            USING (session_id)
                                          INNER JOIN sensemyfeup_raw.session_offsets 
                                            USING (session_id)
                                       WHERE (osm.seconds + session_offsets.clock_offset_seconds BETWEEN segments.seconds_start AND seconds_end) AND
                                            segments.travelmode = 'car' OR segments.travelmode =  'bus'
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_osm_cars, file="df_points_edge_osm_cars2.Rda")
```

### SESSION per edge. 
Getting the sessions filtered by only segments with travelmode = car or bus.

```{r}

df_sessions_edge_osm_cars <- data.frame()
df_sessions_edge_osm_cars <- dbGetQuery(con, "SELECT way_id, COUNT(DISTINCT session_id) AS sessions
                                   FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_travelmode as segments 
                                            USING (session_id)
                                          INNER JOIN sensemyfeup_raw.session_offsets 
                                            USING (session_id)
                                       WHERE (osm.seconds + session_offsets.clock_offset_seconds BETWEEN segments.seconds_start AND seconds_end) AND
                                            segments.travelmode = 'car' OR segments.travelmode =  'bus'
                                       GROUP BY way_id
                                       ORDER BY sessions DESC")

save(df_sessions_edge_osm_cars, file="df_sessions_edge_osm_cars.Rda")
```


## Plots edges and LOADING
### POINTS 
```{r echo=FALSE, eval=FALSE}
# REFERENCE BY POINTS
#Loading df about to use.

load("df_points_edge_osm_cars.Rda")
df_osm_edge <- df_points_edge_osm_cars

df_osm_edge <- df_points_edge_cars_april16

```

#### Histograms
```{r}
p1 <- subset(df_osm_edge, points < mean(points)) 
summary(p1)

ggplot(p1, aes(points)) + geom_histogram(binwidth = 10) +
  xlim(0,1000)


```

#### ECDF 
```{r echo=FALSE, eval=FALSE}
# Listing and obtaining ECDF pf the points.

prt <- 1
set.seed(4218)
sp <- sample(1:nrow(df_osm_edge), as.integer(prt * nrow(df_osm_edge)))

list_osm_edge <- df_osm_edge[sp, 1]
fun.ecdf <- ecdf(df_osm_edge$points)

my.ecdf <- fun.ecdf(sort(df_osm_edge$points))

my_ecdf_df <- data.frame(my.ecdf)

my_ecdf_df$points <- sort(df_osm_edge$points)
tail(my_ecdf_df, 400)

P <- ecdf(subset(df_osm_edge, points < quantile(df_osm_edge$points,.9))$points)

#grid()
```

```{r eval= TRUE}
plot(P, log="x", xlim=c(1, max(df_osm_edge$points)))

```

### SESSIONS 
```{r echo=FALSE, eval=FALSE}
#REFERENCE BY SESSIONS

load("df_sessions_edge_osm_cars.Rda")
df_osm_edge <- df_sessions_edge_osm_cars
```


```{r}
p1 <- subset(df_osm_edge, sessions < mean(sessions)) 
summary()

ggplot(p1, aes(sessions)) + geom_histogram(binwidth=1) +
  scale_x_continuous(breaks = seq(0,8,1))

```

####ECDF
```{r}
list_osm_edge <- df_osm_edge[, 1]

fun.ecdf <- ecdf(df_osm_edge$sessions)

my.ecdf <- fun.ecdf(sort(df_osm_edge$sessions))

my_ecdf_df <- data.frame(my.ecdf)

my_ecdf_df$sessions <- sort(df_osm_edge$sessions)

tail(my_ecdf_df, 400)

P <- ecdf(df_osm_edge$sessions)
```

```{r eval= TRUE}
plot(P, log="x", xlim=c(1, max(df_osm_edge$sessions)))
```

## Map points/sessions per edge.

```{r echo=FALSE, eval=FALSE}
# Map color parameter. 

# Require to be connected to con_osm
# Going through every point of the df selected 

# Defining limits to every color. 
# More than superior is red.
# More than medio and less than superior is yellow.
# Less than medio is green.     

#Points
superior <- quantile(df_osm_edge$points, 0.9)
medio <- quantile(df_osm_edge$points, 0.7)
low <- quantile(df_osm_edge$points, 0.5) 

#Sessions 
#superior <- quantile(df_osm_edge$sessions, 0.9)
#medio <- quantile(df_osm_edge$sessions, 0.7)
#low <- quantile(df_osm_edge$sessions, 0.5) 
```

### Mapping 
```{r echo=FALSE, eval=FALSE}
# Creating the empty map of Porto

m <- leaflet() %>% setView(lng=-8.61419, lat=41.16311, zoom = 13)
m <- addTiles(m) 
m <- addProviderTiles(m, "CartoDB.Positron")

counter <- 1

for(way_id in list_osm_edge) {
  
  df_way_id <- dbGetQuery(con_osm, paste0("SELECT st_astext(st_transform(way, 4326)) AS line FROM planet_osm_line WHERE planet_osm_line.osm_id = ", way_id))
  
  line <- df_way_id$line
  line <- as.character(line)
  
  line <- unlist(strsplit(line, split='(', fixed=TRUE))[2]
  line <- substr(line, 1, nchar(line) - 1)
  
  parsed_line <- strsplit(line, ",")
  
  lons <- c()
  lats <- c()
  
  if(length(parsed_line) != 0) {
    
    #Defining lons and lats as variables to use later.

    for(coord in parsed_line[[1]]) {
      
      lon <- unlist(strsplit(coord, split=' ', fixed=TRUE))[1]
      lat <- unlist(strsplit(coord, split=lon, fixed=TRUE))[2]
      lat <- substr(lat, 2, nchar(lat))
      
      lon <- as.numeric(lon)
      lat <- as.double(lat)
      
      lons <- c(lons, lon)
      lats <- c(lats, lat)
      
    }
    
    # Deciding the color of the point.

    if(df_osm_edge[counter, 2] > superior) {
          
          m <- addPolylines(m, lons, lats, color='red')
          
        } else if (df_osm_edge[counter, 2] >= medio && df_osm_edge[counter, 2] <= superior) {
          
          m <- addPolylines(m, lons, lats, color='yellow')
          
        } else if (df_osm_edge[counter, 2] >= low && df_osm_edge[counter, 2] <= medio) {
          
          m <- addPolylines(m, lons, lats, color='green')
          
        }
        
        counter <- counter + 1 
        
      }
      
      #print(line)
      
    }

```

```{r eval= TRUE}
# Showing Map.
m 
```

## Writing the results

```{r echo=FALSE}
#Points
write.xlsx(df_osm_edge, file="df_points_edge_osm.xlsx")

#Sessions
write.xlsx(df_osm_edge, file="df_sessions_edge_osm.xlsx")

```


# Disconnecting Databases.
```{r echo=FALSE, eval=FALSE}
dbDisconnect(con_osm)
dbDisconnect(con)

```




# PART II: only data of April 2016.

## Getting data from database.

### Points April16
All
```{r}
df_points_edge_cars_april16 <- data.frame()
df_points_edge_cars_april16 <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM segments_vehicules_april as segments
                                       INNER JOIN sensemyfeup_raw.gslocation as gs
                                         USING (session_id)
                                       INNER JOIN sensemyfeup_raw.osmlocation as osm
                                         USING (session_id)
                                       WHERE (osm.seconds between segments.seconds_start AND segments.seconds_end) AND
gs.seconds = osm.seconds AND gs.millis = osm.millis AND 
(alt != 0  OR  track != 0)
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16, file="df_points_edge_cars_april16.Rda")
```

Day time
```{r}
df_points_edge_cars_april16_day <- data.frame()
df_points_edge_cars_april16_day <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_vehicules_april as segments 
                                            USING (session_id)
INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                                       WHERE (osm.seconds  BETWEEN segments.seconds_start AND seconds_end) AND 
(alt != 0  OR  track != 0)  AND
EXTRACT(HOUR from time_start) BETWEEN 7 and 20
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16_day, file="df_points_edge_cars_april16_day.Rda")
```

Night time 
```{r}
df_points_edge_cars_april16_night <- data.frame()
df_points_edge_cars_april16_night <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_vehicules_april as segments 
                                            USING (session_id)
INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                                       WHERE (osm.seconds  BETWEEN segments.seconds_start AND seconds_end) AND 
EXTRACT(HOUR from time_start) > 20 OR EXTRACT(HOUR from time_start) < 7 AND 
(alt != 0  OR  track != 0)
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16_night, file="df_points_edge_cars_april16_night.Rda")
```


### Just first session, April16 
```{r echo=FALSE, eval=FALSE}
df_first_session_cars_april16 <- data.frame()
df_first_session_cars_april16 <- dbGetQuery(con, "SELECT distinct on (session_id) as session_id, time_start,time_end
                                          FROM danielasocas.segments_travelmode AS segments
                                          WHERE (segments.travelmode = 'car' OR segments.travelmode =  'bus') AND
                                          EXTRACT(MONTH from time_start) = 4 AND EXTRACT(YEAR from time_start) = 2016
                                          ORDER BY session_id, time_start")

save(df_first_session_cars_april16, file="df_first_session_cars_april16.Rda")
```

### Way id time intersession 

```{r}

#Not anymore

df_intersession_april_gs <- data.frame()
df_intersession_april_gs <- dbGetQuery(con, "SELECT session_id, way_id, MIN(osm.seconds) as seconds
                                                  FROM sensemyfeup_raw.osmlocation AS osm
                                                  INNER JOIN danielasocas.segments_vehicules_april AS segments
	                                                  USING (session_id)
                                                  INNER JOIN sensemyfeup_raw.session_offsets 
	                                                  USING (session_id)
INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                                                  WHERE (osm.seconds + session_offsets.clock_offset_seconds BETWEEN segments.seconds_start AND seconds_end) AND
(alt != 0  OR  track != 0)
                                                  GROUP BY way_id, session_id
                                                  ORDER BY way_id")

save(df_intersession_april_gs, file="df_intersession_april_gs.Rda")

```

### Speed

```{r}

df_speed <- data.frame()
df_speed <- dbGetQuery(con, "SELECT session_id, segment_id, way_id, osm.seconds, osm.lat, osm.lon, gs.speed
                            FROM segments_vehicules_april as segments
                              INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                              INNER JOIN sensemyfeup_raw.osmlocation as osm
                                USING (session_id)
                            WHERE (osm.seconds between segments.seconds_start AND segments.seconds_end) AND
gs.seconds = osm.seconds AND gs.millis = osm.millis AND 
(alt != 0  OR  track != 0)")

save(df_speed, file="df_speed.Rda")

```







