---
title: "Pre Analysis - top edges."
output: html_notebook
---

<!-- The datasets to make this notebook are in SenseMyFEUP.Rmd
     The rest of the code is divided: 
          General information: April2016.Rmd 
          Timeseries code: timeseries.R
          
    df_speed_top it is on preanalysis_course.Rmd
-->

# Introduction 

This report explores the top 9 edges with more sessions from April 2016.
The goal is to understand this data for later modeling and feature decisions. 

#### Index: 

  * Map.
  * Number of sessions.
  * Parking.
  * Time discretization. 
  * Speed and intersession time.
  * State discretization. 
  
--------------------------------------------------------------------------------
# Setting dataframe with top edges in April 2016(vehicules).  
 Filters: 
  * No parking edges.
  * Week days.
  * Hour between 8hr and 20h.
  
Rua Eduardo Silva is divided inthe intersection for better analysis. 

# Map of top points 

```{r echo=FALSE}
#m_top9

df_sessions_edges_april16pt %>% 
  subset(way_id %in% df_superhotedges_april16pt$way_id)  %>% 
  map()
```

#
## Low speed and Parkings

```{r}
df_superhotedges_april16pt %>% 
  group_by(way_id, session_id,segment_id) %>% 
  summarise(med_tr =median(track),iqr = IQR(track), mean= mean(speed*18/5),
            median =  median(speed*18/5),q75=  quantile(speed*18/5, 0.75)) %>% 
  merge(y= df_osm_edge_ids) %>% 
  ggplot(aes(x=iqr, fill = median)) +
  geom_histogram(binwidth = 20) + 
  facet_wrap(~type~address)
```


```{r echo=FALSE}
# Frei 
df_superhotedges_april16pt %>% 
  filter(way_id == 37141967) %>% 
  group_by(session_id, segment_id) %>% 
  filter( IQR(track) >100, lon < -8.5981 , lon > -8.5944   ) 

# Others. 
df_superhotedges_april16pt %>% 
  filter(way_id == 37141967) %>% 
  group_by(session_id, segment_id) %>% 
  filter( IQR(track) >100, lon < -8.5981 , lon > -8.5944   ) 

  
df_superhotedges_april16pt %>% 
  filter(way_id == 37141967, day(time) < 5) %>% 
  group_by(session_id, segment_id) %>% 
  ggplot(aes(x=factor(session_id), y=track, color= factor(segment_id))) +
  geom_boxplot(outlier.shape = NA) + 
  geom_hline(aes(yintercept = 90)) + 
  geom_hline(aes(yintercept = 180)) +
  geom_hline(aes(yintercept = 270)) +
  theme(legend.position="none")+ 
  ggtitle(paste("Median of the track by session and segments 37141967", sep = " " )) + 
  xlab("Grades(º)") +
  ylab("Density") 

```

# Time discretization.

# Number of sessions per edge. 
```{r}
df_superhotedges_april16pt %>% 
  group_by(h = hour(time)) %>%  
  summarise(n = n_distinct(session_id))  %>% 
  ggplot(aes(x= h, y = n)) +
  geom_line() + 
  scale_x_continuous(breaks = seq(8,20,1)) +
  xlab("Hour") + 
  ylab("# sessions")
```


```{r echo=FALSE}
dfm_base %>% 
  ggplot(aes(x= hour, fill = course)) + 
    geom_histogram(binwidth = 1) +
  facet_wrap(~heading)+
  scale_x_continuous(breaks = seq(8,20,2)) + 
  ggtitle("Sessions by hour and course for everyday(WW)")
```

```{r echo=FALSE}
dfm_base %>% 
  ggplot(aes(x= period, fill = period)) + 
    geom_histogram(binwidth = 1, stat = "count") +
  facet_wrap(~day)+
  ggtitle("Sessions by hour and course for everyday(WW)")
```

```{r}
dfm_base %>% 
  ggplot(aes(x= period, fill = factor(hour))) + 
    geom_histogram(binwidth = 1, stat = "count") 

```


```{r echo=FALSE, message=FALSE}
day_hist1 <-  function(index) { dfm_base %>% 
  filter( day == index ) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x= factor(hour))) +
  geom_histogram(binwidth = 1, aes(group=address, color=address), stat = "count") +
  facet_wrap(~heading)  +
  ggtitle(paste("NºSessions by hour, edge and course on April:", index, sep = " " )) + 
  xlab("Hour") +
  ylab("Count") 
}

unique(dfm_base$day) %>% 
  sort() %>% 
lapply(FUN = day_hist1)
```

# Speed & intersession times

## Boxplot intersession time  

```{r echo=FALSE}

df_super_intersession_april16pt <- df_superhotedges_april16pt %>% 
  group_by(way_id, session_id, heading) %>% 
  summarise(seconds = min(seconds))

df_super_intersession_april16pt$time <- as.POSIXct(df_super_intersession_april16pt$seconds, origin="1970-01-01")

#df_super_intersession_april16pt <- 
  
  df_super_intersession_april16pt  %>%
  arrange(way_id, time, heading) %>% 
  mutate(intersession_time = c(0,as.numeric(diff(time), units="mins")))

```


```{r echo=FALSE}
#Boxplots all 
df_super_intersession_april16pt %>%
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(factor(address), intersession_time)) +
  geom_boxplot(aes(color= heading )) +
  coord_cartesian(ylim= c(0,10))+
  ggtitle(label = "Intersession time top 12 edges") + 
  xlab("") + 
  ylab("Time (min)") +
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))
```

```{r echo=FALSE}
lapply(1:8, FUN = newspeed_avg_h)

```

```{r echo=FALSE}
lapply(1:8, FUN = speed_s)
```

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  group_by(way_id,heading, hour = hour(time)) %>%
  summarise(avg_speed = mean(speed*18/5), sd = sd(speed)) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x=avg_speed, y=sd, color = hour)) +
  geom_point() +
  geom_smooth(aes(colour = avg_speed, fill = avg_speed)) +
  facet_wrap(~ address~heading) + 
  ggtitle("By hour avg. speed April16")
```

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( wday(time) != 1 , wday(time) != 7 ) %>% 
  group_by(way_id,heading,day = day(time), hour = hour(time)) %>% 
  summarise(avg_speed = mean(speed*18/5), sd = sd(speed)) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x=avg_speed, y=sd, color = hour)) +
  geom_point() +
  geom_smooth(aes(colour = avg_speed, fill = avg_speed)) +
  facet_wrap(~ address ~ heading) + 
  ggtitle("By day and hour avg.speed")
```




# State discretization.

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  group_by(session_id) %>% 
  summarise(speed = mean(speed*18/5), n= n()) %>% 
  mutate(flow = case_when(.$speed <= 18 ~ "congestion", .$speed > 18 & .$speed < 30 ~ "sync", .$speed >= 30 ~"free" )) %>% 
  ggplot(aes(x=n, y = speed, color = flow)) + 
  geom_point() + 
  coord_cartesian(xlim = c(0,400)) + 
  xlab("Distinct sessions") + 
  ylab("Speed (km/h)") + 
  ggtitle("State discretization top 7 edges.")
```

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  group_by(session_id) %>% 
  summarise(speed = mean(speed*18/5)) %>% 
  mutate(flow = case_when(.$speed <= 18 ~ "congestion", .$speed > 18 & .$speed < 30 ~ "sync", .$speed >= 30 ~"free" )) %>% 
  ggplot(aes(x=speed, fill = flow)) + 
  geom_histogram(binwidth = 2) 
```



