---
title: "Sense My FEUP - Mobility forecasting based on GPS location tracking"
author: "Daniela S. Gil"
date: 25-08-2017
output: html_notebook
---

# Introduction

This Notebook concentrates to be the base file of the project. 
Meaning to have all the basic steps for his execution and it is 
not suppose to run all, it is just for guidance and order.

#### Index:

  * Packages to install.
  * Connection to the database.
  * Mapping. 
  * Getting data from database.
    +  Points.
    +  Sessions.
    +  All info. 
    +  SenseMyCity, SenseMood

--------------------------------------------------------------------------------

# PART I: system setup.  

## Package installation. 
```{r echo=FALSE, eval=FALSE}
#Basics. 

if(! "ggplot2" %in% rownames(installed.packages())){
  install.packages("ggplot2")
}

if(! "dplyr" %in% rownames(installed.packages())){
  install.packages("dplyr")
}

if(! "tidyr" %in% rownames(installed.packages())){
  install.packages("tidyr")
}

#if(! "xlsx" %in% rownames(installed.packages())){
#  install.packages("xlsx")
#}

if(! "magrittr" %in% rownames(installed.packages())){
  install.packages("magrittr")
}

if(! "gridExtra" %in% rownames(installed.packages())){
  install.packages("gridExtra")
}

if(! "plotly" %in% rownames(installed.packages())){
  install.packages("plotly")
}
if(! "corrgram" %in% rownames(installed.packages())){
  install.packages("corrgram")
}
if(! "reshape2" %in% rownames(installed.packages())){
  install.packages("reshape2")
}

library(gridExtra)
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(corrgram)
library(reshape2)
library(magrittr)
#library(xlsx)
#library(magrittr)
#library(lattice)
#library(latticeExtra)
```

```{r echo=FALSE, eval=FALSE}
#connection to database.

if(! "DBI" %in% rownames(installed.packages())){
  install.packages("DBI")
}
if(! "RPostgreSQL" %in% rownames(installed.packages())){
  install.packages("RPostgreSQL")
}

library(DBI)
library(RPostgreSQL)
```

```{r echo=FALSE, eval=FALSE}
#Mapping. 

if(! "leaflet" %in% rownames(installed.packages())){
  install.packages("leaflet",dependencies = TRUE)
}

#library(htmlwidgets)
#library(mapview)
library(leaflet)
```

```{r echo=FALSE, eval=FALSE}
#Time management.

if(! "lubridate" %in% rownames(installed.packages())){
  install.packages("lubridate")
}

library(lubridate)
```

```{r echo=FALSE, eval=FALSE}
#Write Excel

install.packages("rJava")

library(rJava)
library(xlsxjars)
library(xlsx)

```

```{r echo=FALSE, eval=FALSE}
#Modeling.

if(! "DMwR" %in% rownames(installed.packages())){
  install.packages("DMwR")
}
if(! "rpart" %in% rownames(installed.packages())){
  install.packages("rpart")
}
if(! "rpart.plot" %in% rownames(installed.packages())){
  install.packages("rpart.plot")
}
if(! "Hmisc" %in% rownames(installed.packages())){
  install.packages("Hmisc")
}
if(! "earth" %in% rownames(installed.packages())){
  install.packages("earth")
}
if(! "randomForest" %in% rownames(installed.packages())){
  install.packages("randomForest")
}
if(! "pROC" %in% rownames(installed.packages())){
  install.packages("pROC")
}
if(! "caret" %in% rownames(installed.packages())){
  install.packages("caret")
}
if(! "rattle" %in% rownames(installed.packages())){
  install.packages("rattle")
}

library(DMwR)
library(rpart)
library(rpart.plot)
library(Hmisc)
library(earth)
library(randomForest)
library(pROC)
library(caret)
library(rattle)
```

## Connecting the database.
```{r echo=FALSE, eval=FALSE}
# Connecting Sense My FEUP - DS

drv <- dbDriver("PostgreSQL")
#pwd <- password DS 

#Database to connect 
db <- "sensemyfeup"
#db <- "sensemycity"
#db <- "moodsensor"

con <-dbConnect(drv,host='localhost',port='15432',dbname=db,user='danielasocas', 
                password=pwd)
```

```{r echo=FALSE, eval=FALSE}
# Connecting OpenStreetMap - Marcio Fontes
#pw <- password Marcio Fontes

drv <- dbDriver("PostgreSQL")

con_osm <- dbConnect(drv, dbname = "openstreetmap",
                     host = "localhost", port = 15432,
                     user = "marciofontes", password = pw)
```

## Disconnecting Databases.
```{r echo=FALSE, eval=FALSE}
dbDisconnect(con_osm)
dbDisconnect(con)
```



# Part II:  First attempts.

This is the first attempt to understand the data. This is replaced later with 
better coding but its left here to compare first attempts and backup. Mapping it
is then separated in functions to easier approach. 
Specially chunks to get data from database and plot:
  
  * Distance by sessions
  * Getting data per edge
    + Number of points per edge 
    + Number of sessions per edge
  * Mapping 
  * ECDF 
  
## Distance by sessions 
```{r echo= FALSE}
df_segments_length <- data.frame()
df_segments_length_tm <- dbGetQuery(con, "SELECT session_id, SUM(length) as distance, travelmode
                                      FROM danielasocas.segments_travelmode as segments 
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'
                                      GROUP BY session_id, travelmode")

df_segments_length <- dbGetQuery(con, "SELECT session_id, SUM(length) as distance
                                      FROM danielasocas.segments_travelmode as segments 
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'
                                      GROUP BY session_id")

```

```{r echo= FALSE}
df_segments <- subset(df_segments_length, distance < IQR(distance,0.75) )

ggplot(df_segments, aes(distance)) +
  geom_histogram(bins = 10)

ggplot(df_segments_length, aes(travelmode, distance, color = travelmode)) +
  geom_boxplot()

ggplot(df_segments, aes(distance, color = travelmode)) +
  geom_histogram(binwidth = 1)

ggplot(df_segments, aes(travelmode, distance, color = travelmode)) +
  geom_boxplot()
```


```{r echo= FALSE}
df_segments_timestamp <- data.frame()
df_segments_timestamp <- dbGetQuery(con, "SELECT session_id, time_start,time_end, length, travelmode
                                      FROM danielasocas.segments_travelmode as segments
                                      WHERE segments.travelmode = 'car' OR segments.travelmode = 'bus'")
```


## Getting data per edge.
### POINTS per edge. 
Getting the points filtered by only segments with travelmode = car or bus.

```{r echo=FALSE, eval=FALSE}
#Obtaining the points through the session and the seconds of the whole segment.  

df_points_edge_osm_cars <- data.frame()
df_points_edge_osm_cars <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_travelmode as segments 
                                            USING (session_id)
                                          INNER JOIN sensemyfeup_raw.session_offsets 
                                            USING (session_id)
                                       WHERE (osm.seconds + session_offsets.clock_offset_seconds BETWEEN segments.seconds_start AND seconds_end) AND
                                            segments.travelmode = 'car' OR segments.travelmode =  'bus'
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_osm_cars, file="df_points_edge_osm_cars.Rda")
```

### SESSION per edge. 
Getting the sessions filtered by only segments with travelmode = car or bus.

```{r}

df_sessions_edge_osm_cars <- data.frame()
df_sessions_edge_osm_cars <- dbGetQuery(con, "SELECT way_id, COUNT(DISTINCT session_id) AS sessions
                                   FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_travelmode as segments 
                                            USING (session_id)
                                          INNER JOIN sensemyfeup_raw.session_offsets 
                                            USING (session_id)
                                       WHERE (osm.seconds + session_offsets.clock_offset_seconds BETWEEN segments.seconds_start AND seconds_end) AND
                                            segments.travelmode = 'car' OR segments.travelmode =  'bus'
                                       GROUP BY way_id
                                       ORDER BY sessions DESC")

save(df_sessions_edge_osm_cars, file="df_sessions_edge_osm_cars.Rda")
```




## Plots edges and LOADING
### POINTS 
```{r echo=FALSE, eval=FALSE}
# NECESSARY FOR MAPPING
# REFERENCE BY POINTS
#Loading df about to use.

load("df_points_edge_osm_cars.Rda")
df_osm_edge <- df_points_edge_osm_cars

df_osm_edge <- df_points_edge_cars_april16

```

#### Histograms
```{r}
p1 <- subset(df_osm_edge, points < mean(points)) 
summary(p1)

ggplot(p1, aes(points)) + geom_histogram(binwidth = 10) +
  xlim(0,1000)
```

#### ECDF 
```{r echo=FALSE, eval=FALSE}
# Listing and obtaining ECDF pf the points.

prt <- 1
set.seed(4218)
sp <- sample(1:nrow(df_osm_edge), as.integer(prt * nrow(df_osm_edge)))

list_osm_edge <- df_osm_edge[, 1]
fun.ecdf <- ecdf(df_osm_edge$points)

my.ecdf <- fun.ecdf(sort(df_osm_edge$points))

my_ecdf_df <- data.frame(my.ecdf)

my_ecdf_df$points <- sort(df_osm_edge$points)
tail(my_ecdf_df, 400)

P <- ecdf(subset(df_osm_edge, points < quantile(df_osm_edge$points,.9))$points)

#grid()
```

```{r eval= TRUE}
plot(P, log="x", xlim=c(1, max(df_osm_edge$points)))
```

### SESSIONS 
```{r echo=FALSE, eval=FALSE}
#REFERENCE BY SESSIONS

load("df_sessions_edge_osm_cars.Rda")
df_osm_edge <- df_sessions_edge_osm_cars
```


```{r}
p1 <- subset(df_osm_edge, sessions < mean(sessions)) 
summary()

ggplot(p1, aes(sessions)) + geom_histogram(binwidth=1) +
  scale_x_continuous(breaks = seq(0,8,1))

```

####ECDF
```{r}
list_osm_edge <- df_osm_edge[, 1]

fun.ecdf <- ecdf(df_osm_edge$sessions)

my.ecdf <- fun.ecdf(sort(df_osm_edge$sessions))

my_ecdf_df <- data.frame(my.ecdf)

my_ecdf_df$sessions <- sort(df_osm_edge$sessions)

tail(my_ecdf_df, 400)

P <- ecdf(df_osm_edge$sessions)
```

```{r eval= TRUE}
plot(P, log="x", xlim=c(1, max(df_osm_edge$sessions)))
```

## Map points/sessions per edge.

```{r echo=FALSE, eval=FALSE}
# Map color parameter. 

# Require to be connected to con_osm
# Going through every point of the df selected 

# Defining limits to every color. 
# More than superior is red.
# More than medio and less than superior is yellow.
# Less than medio is green.     

#Points
superior <- quantile(df_osm_edge$points, 0.9)
medio <- quantile(df_osm_edge$points, 0.7)
low <- quantile(df_osm_edge$points, 0.5) 

#Sessions 
#superior <- quantile(df_osm_edge$sessions, 0.9)
#medio <- quantile(df_osm_edge$sessions, 0.7)
#low <- quantile(df_osm_edge$sessions, 0.5) 
```

### Mapping 
```{r echo=FALSE, eval=FALSE}
# Creating the empty map of Porto

m <- leaflet() %>% setView(lng=-8.61419, lat=41.16311, zoom = 13)
m <- addTiles(m) 
m <- addProviderTiles(m, "CartoDB.Positron")

counter <- 1

for(way_id in list_osm_edge) {
  
  df_way_id <- dbGetQuery(con_osm, paste0("SELECT st_astext(st_transform(way, 4326)) AS line FROM planet_osm_line WHERE planet_osm_line.osm_id = ", way_id))
  
  line <- df_way_id$line
  line <- as.character(line)
  
  line <- unlist(strsplit(line, split='(', fixed=TRUE))[2]
  line <- substr(line, 1, nchar(line) - 1)
  
  parsed_line <- strsplit(line, ",")
  
  lons <- c()
  lats <- c()
  
  print(length(parsed_line))
  if(length(parsed_line) != 0) {
    
    #Defining lons and lats as variables to use later.

    for(coord in parsed_line[[1]]) {
      
      lon <- unlist(strsplit(coord, split=' ', fixed=TRUE))[1]
      lat <- unlist(strsplit(coord, split=lon, fixed=TRUE))[2]
      lat <- substr(lat, 2, nchar(lat))
      
      lon <- as.numeric(lon)
      lat <- as.double(lat)
      
      lons <- c(lons, lon)
      lats <- c(lats, lat)
      
    }
    
    # Deciding the color of the point.

    if(df_osm_edge[counter, 2] > superior) {
          
          m <- addPolylines(m, lons, lats, color='red')
          
        } else if (df_osm_edge[counter, 2] >= medio && df_osm_edge[counter, 2] <= superior) {
          
          m <- addPolylines(m, lons, lats, color='yellow')
          
        } else if (df_osm_edge[counter, 2] >= low && df_osm_edge[counter, 2] <= medio) {
          
          m <- addPolylines(m, lons, lats, color='green')
          
        }
        
        counter <- counter + 1 
      }
      
      #print(line)
      
    }

```

```{r eval= TRUE}
# Showing Map.
m 
```

## Writing the results

```{r echo=FALSE}
#Points
write.xlsx(df_osm_edg, file="df_points_edge_osm.xlsx")

#Sessions
write.xlsx(df_osm_edge, file="df_sessions_edge_osm.xlsx")

```




# PART III: only data of April 2016.

 Getting data from database, exclusively for April 2016.
##Points
All
```{r}
df_points_edge_cars_april16 <- data.frame()
df_points_edge_cars_april16 <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM segments_vehicules_april as segments
                                       INNER JOIN sensemyfeup_raw.gslocation as gs
                                         USING (session_id)
                                       INNER JOIN sensemyfeup_raw.osmlocation as osm
                                         USING (session_id)
                                       WHERE (osm.seconds between segments.seconds_start AND segments.seconds_end) AND
gs.seconds = osm.seconds AND gs.millis = osm.millis AND 
(alt != 0  OR  track != 0)
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16, file="df_points_edge_cars_april16.Rda")
```

Day time
```{r}
df_points_edge_cars_april16_day <- data.frame()
df_points_edge_cars_april16_day <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_vehicules_april as segments 
                                            USING (session_id)
INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                                       WHERE (osm.seconds  BETWEEN segments.seconds_start AND seconds_end) AND 
(alt != 0  OR  track != 0)  AND
EXTRACT(HOUR from time_start) BETWEEN 8 and 20
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16_day, file="df_points_edge_cars_april16_day.Rda")
```

Night time 
```{r}
df_points_edge_cars_april16_night <- data.frame()
df_points_edge_cars_april16_night <- dbGetQuery(con, "SELECT way_id, COUNT(*) AS points
                                       FROM sensemyfeup_raw.osmlocation as osm 
                                          INNER JOIN danielasocas.segments_vehicules_april as segments 
                                            USING (session_id)
INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                                       WHERE (osm.seconds  BETWEEN segments.seconds_start AND seconds_end) AND 
EXTRACT(HOUR from time_start) > 20 OR EXTRACT(HOUR from time_start) < 8 AND 
(alt != 0  OR  track != 0)
                                       GROUP BY way_id
                                       ORDER BY points DESC")

save(df_points_edge_cars_april16_night, file="df_points_edge_cars_april16_night.Rda")
```

## Speed
Data used for most analysis.
* Speed: data used for April16, all edges and relevant features.
* Top speed: data for preanalysis top edges, includes track.

It is in To do list change the name of this because it is not only about speed. 
It is the main dataframe of this project. 

```{r}

df_speed <- data.frame()
df_speed <- dbGetQuery(con, "SELECT session_id, segment_id, way_id, osm.seconds, osm.lat, osm.lon, gs.speed
                            FROM segments_vehicules_april as segments
                              INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                              INNER JOIN sensemyfeup_raw.osmlocation as osm
                                USING (session_id)
                            WHERE (osm.seconds between segments.seconds_start AND segments.seconds_end) AND
gs.seconds = osm.seconds AND gs.millis = osm.millis AND 
(alt != 0  OR  track != 0)")

save(df_speed, file="df_speed.Rda")

```

```{r}
df_speed_top <- data.frame()
df_speed_top <- dbGetQuery(con, "SELECT session_id, segment_id, way_id, osm.seconds, osm.lat, osm.lon, gs.speed, gs.track
                            FROM segments_vehicules_april as segments
                              INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                              INNER JOIN sensemyfeup_raw.osmlocation as osm
                                USING (session_id)
                            WHERE (osm.seconds between segments.seconds_start AND segments.seconds_end) AND
gs.seconds = osm.seconds AND gs.millis = osm.millis AND 
(alt != 0  OR  track != 0) AND (way_id = 35078500 OR way_id = 35145521 OR way_id = 316459169 OR 
way_id = 37141967 OR way_id = 63192525 OR way_id = 106334851 OR way_id = 37141970 OR way_id = 36979228) ")

save(df_speed_top, file="df_speed_top.Rda")
```

## Sessions all 
Number of sessions for all transport modes in April and in general. 

```{r}
df_sessions_allmodes_april16pt <- data.frame()
df_sessions_allmodes_april16pt <- dbGetQuery(con, "SELECT *
                            FROM danielasocas.sessions_allmodes_april16pt")

save(df_sessions_allmodes_april16pt, file="df_sessions_allmodes_april16pt.Rda")
```

```{r}
df_sessions_allmodes <- data.frame()
df_sessions_allmodes <- dbGetQuery(con, "SELECT *
                            FROM danielasocas.sessions_allmodes")

save(df_sessions_allmodes, file="df_sessions_allmodes.Rda")
```


# Part IV 

## Sense my FEUP  
```{r}
df_smf_all <- dbGetQuery(con, "SELECT session_id, daily_user_id, osm.seconds, osm.lat, osm.lon
                            FROM sensemyfeup.sessions 
                              INNER JOIN sensemyfeup_raw.gslocation as gs
                                USING (session_id)
                              INNER JOIN sensemyfeup_raw.osmlocation as osm
                                USING (session_id, seconds, millis)
                            WHERE (osm.seconds > sessions.seconds_start) AND
(alt != 0  OR  track != 0)")

save(df_smf_all, file="df_smf_all.Rda")
```

## Sense my City  
```{r}
df_smc_all <- dbGetQuery(con, "SELECT session_id, daily_user_id, loc.seconds, loc.lat, loc.lon
                            FROM sensemycity_raw.session
                              INNER JOIN sensemycity_raw.gslocation as gs
                                USING (session_id)
                              INNER JOIN sensemycity_raw.location as loc
                                USING (session_id, seconds, millis)
                            WHERE (loc.seconds > session.start_time) AND
(gs.alt != 0  OR  gs.track != 0)")

save(df_smc_all, file="df_smc_all.Rda")
```

## Sense my Mood  
```{r}
df_smm_all <- dbGetQuery(con, "SELECT session_id, daily_user_id, loc.seconds, loc.lat, loc.lon
                            FROM public.session
                              INNER JOIN public.gslocation as gs
                                USING (session_id)
                              INNER JOIN public.location as loc
                                USING (session_id, seconds, millis)
                            WHERE (loc.seconds > session.start_time) AND
(gs.alt != 0  OR  gs.track != 0)")

save(df_smm_all, file="df_smm_all.Rda")
```
