---
title: "Correlation between timeseries"
date: 26-09-2017
output: 
  html_notebook:
  toc: true
  
---

# Introduction 

Correlation analisis between edges divided by course.
  * 1 = going to FEUP
  * 2 = leaving FEUP 

--------------------------------------------------------------------------------

## Course to FEUP 

The first thing we can appreciate its the strong correlation between adjacent 
streets. Specially in the following streets:
  * Henrique Sousa 1 and 2 
  * Manuel Pereira with Roberto Frias 1 
 

```{r echo=FALSE, warning= FALSE}
df_superhotedges_april16pt %>% 
inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(ends_with("1")) %>% 
  corrgram(main = "To Feup (1)", upper.panel=panel.pts, lower.panel=panel.cor)

```

## Course from FEUP 

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(ends_with("2")) %>% 
  corrgram(main = "From Feup (2)", upper.panel=panel.pts, lower.panel=panel.cor)

```

Happens in both ways:
  * Strong correlation between Henrique Sousa Reis 1 e 2. 

## Opposite course

```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(contains("Eduardo")) %>% 
  corrgram(main = "Eduardo Silva", upper.panel=panel.pts, lower.panel=panel.cor)

```


```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(contains("Frias")) %>% 
  corrgram(main = "Roberto Frias", upper.panel=panel.pts, lower.panel=panel.cor)

```


# By hour 

## Mannha all
```{r echo=FALSE, message=FALSE, warning=FALSE}
df_superhotedges_april16pt %>% 
  filter( hour(time) == 9 | hour(time) == 10) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  corrgram(main = "All morning", upper.panel=panel.pts, lower.panel=panel.cor)
```


## Night 
```{r echo=FALSE, message=FALSE, warning=FALSE}
df_superhotedges_april16pt %>% 
  filter( hour(time) == 18 | hour(time) == 19) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  corrgram(main = "All night", upper.panel=panel.pts, lower.panel=panel.cor)
```


## Other way 
  
  Different than at night, you can see some high speeds at Roberto Frias 2. 
  

```{r, message=FALSE, warning=FALSE, echo=FALSE}
df_st_hotedges_april16pt %>% 
  filter(start == "OUT", end == "FEUP" | start == "FEUP" & end == "FEUP") %>% 
  inner_join(df_superhotedges_april16pt, by = "session_id") %>% 
  filter( hour(time) == 8 | hour(time) == 9) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id,period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address) %>% 
  corrgram(main = "To FEUP-morning", upper.panel=panel.pts, lower.panel=panel.cor)
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( hour(time) == 18 | hour(time) == 19) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(ends_with("1")) %>% 
  corrgram(main= "From FEUP-morning", upper.panel=panel.pts, lower.panel=panel.cor)
```

  * Frei with Roberto Frias 2 
  * Roberto Frias 2 with Placido 
  * Roberto Frias 2 with Roberto Frias 1 
  * Roberto Frias 1 with Manuel Silva 
  * HSR 2 with Eduardo Silva 2

```{r, message=FALSE, warning=FALSE, echo=FALSE}
df_st_hotedges_april16pt %>% 
  filter(start == "FEUP" & end == "OUT" | start == "FEUP" & end == "FEUP") %>% 
  inner_join(df_superhotedges_april16pt, by = "session_id") %>% 
  filter( hour(time) == 18 | hour(time) == 19, course == 2) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id,period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address) %>% 
  corrgram(main = "From FEUP-night", upper.panel=panel.pts, lower.panel=panel.cor) 
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( hour(time) == 18 | hour(time) == 19) %>% 
  inner_join(df_osm_edge_ids, by = "way_id") %>% 
  group_by(address, session_id, course, period = cut(time, breaks="1 hour")) %>%
  summarise(session_speed = mean(speed*18/5)) %>%  
  group_by(address, course, period) %>%
  summarise(avg_speed = mean(session_speed))  %>% 
  dcast(period~address+course) %>% 
  select(ends_with("1")) %>% 
  corrgram(main = "To FEUP-night", upper.panel=panel.pts, lower.panel=panel.cor)
```