---
title: "Timeseries analysis top edges"
output: 
  html_notebook:
  toc: true
---

# Introduction 

Here is a study of timeseries to evaluate the data we have and be able to discretize
the time. 

#### Index: 

  * Speed.
  * Intersession time.
  * Standard deviation. 

--------------------------------------------------------------------------------

```{r echo=FALSE}
source("../rscripts/packages.R")
source("../rscripts/timeseries.R")
```

With this first approach we can see the almost total absence of points  on the weekends. 
Also, very few points at night in all edges. 
Most points are under 50km/h.

```{r echo=FALSE}
lapply(1:9, FUN = tseries_light)
```

#Timeseries 

## Speed Boxplot.

Median is under 5km/h for every hour and percentile 75 under 30km/h.
At 13h is the time with less difference and lowest speed records. 
```{r echo=FALSE}
#Boxplots all 
df_superhotedges_april16pt %>% 
  filter(speed > 0, wday(time) != 1 , wday(time) != 7 ) %>% 
  group_by(hour = hour(time)) %>% 
  plot_ly(x= ~factor(hour), y= ~(speed*18)/5, type = 'box',boxpoints = FALSE) %>% 
  layout(title = paste("Avg speed per hour (WW)" ),
         yaxis = list(title="Speed (km/h)"),
         xaxis = list(title="Hour"))
```


## Boxplot intersession time  

The intersession time median it is not even under 60min for every edge.
Frei's st. shows an intersession time way smaller than the rest. 
```{r echo=FALSE}
#Boxplots all 
df_intersession_april16pt %>%
  subset(way_id %in% df_superhotedges_april16pt$way_id) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(factor(address), intersession_time)) +
  geom_boxplot() +
  coord_cartesian(ylim= c(0,250))+
  ggtitle(label = "Intersession time top edges") + 
  xlab("") + 
  ylab("Time (min)") +
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))
```




## All speed timeseries per way_id

### Avg speed by session and then time.

```{r echo=FALSE, message=FALSE, warning=FALSE}
lapply(1:9,FUN = plot.timeseries) 
```

#### Hour
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot.speed_avg(df_superhotedges_april16pt, hour)
```

#### Week day
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot.speed_avg(df_superhotedges_april16pt, wday)
```

#### Day
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot.speed_avg(df_superhotedges_april16pt, day)
```


###Average of speed of every point.
####Number of sessions 

  * We can see usually a incremental of number of sessions along the week. Being 
  thursdays and fridays the one with most sessions. 
  *The drop in day 25th it is because holiday.

```{r echo=FALSE, message=FALSE}
lapply(1:9, FUN = tseries_nsession)
```

### Average speed
```{r echo=FALSE, message=FALSE}
lapply(1:9, FUN = tseries_speed)
```

###Intersession time
```{r echo=FALSE, message=FALSE, warning=FALSE}
lapply(1:9, FUN = tseries_inter)
```

##Density lines average speed all together
```{r echo=FALSE}
ggplotly(dens)
```
## Standard deviation

### By hour
```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( wday(time) != 1 , wday(time) != 7 ) %>% 
  group_by(way_id,session_id, hour = hour(time), heading) %>%
  summarise(speed_sess = mean(speed*18/5)) %>%
  group_by(way_id, hour, heading) %>%
  summarise(avg_speed = mean(speed_sess), sd = sd(speed_sess)) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x=avg_speed, y=sd, color = heading)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ address) + 
  ggtitle("By hour avg. speed April16")
```

### By day and hour
```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  filter( wday(time) != 1 , wday(time) != 7 ) %>% 
  group_by(way_id,session_id, day = wday(time), hour = hour(time), heading) %>% 
  summarise(speed_sess = mean(speed*18/5)) %>% 
  group_by(way_id, day, hour, heading) %>% 
  summarise(avg_speed = mean(speed_sess), sd= sd(speed_sess)) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x=avg_speed, y=sd, color = heading)) +
  geom_point(alpha = 1/3) +
  geom_smooth() +
  facet_wrap(~ address) + 
  ggtitle("By day and hour avg.speed")
```

### By session 
```{r echo=FALSE}
df_superhotedges_april16pt %>% 
  group_by(way_id, session_id, hour = hour(time), heading) %>% 
  summarise(avg_speed = mean(speed*18/5), sd = sd(speed)) %>% 
  merge(y = df_osm_edge_ids, using = way_id) %>% 
  ggplot(aes(x=avg_speed, y=sd, color = heading)) +
  geom_point(alpha = 1/5) +
  geom_smooth(aes(colour = heading, fill = heading)) +
  facet_wrap(~address) + 
  ggtitle("By session per hour avg. speed April16")
```

<!--
## All speed timeseries per way_id
Average of speed of every point.

```{r echo=FALSE, message=FALSE}
lapply(1:9, FUN = tseries)
```
-->




